Study of the morphofunctional signal of captivity on the microanatomy of a tarsal bone, the talus, in a wild ungulate, the wild boar (\textit{Sus scrofa})

### 2025-04-13 article.tex
recent/final Mesol

The SVM average cross-validation score was 93.3\% and the accuracy on the final test was 84.6\%. Mesolithic specimens ARC-12990, J234, EG235, G236, C235 and C238 were classified as having microanatomy associated with captivity. Specimens E143-7, E236, G238, K135, M212, ARC-16242, RAN-270, RAN-397, RAN-398, RAN-460, RAN-545, RAN-706 were classified as having microanatomy associated with life in the wild.

@@

%verif si je la mets \subsection{Limitations}  
%As previously mentioned, the actualistic referential framework used here has several limitations. The potential reduction in rooting behavior due to supplemental feeding in the Chambord and Compiègne wild boars may have altered their locomotion. Notably, the control population already occupies a relatively restricted space compared to the Chambord and Compiègne specimens. Moreover, classifying this fenced-forest population as "wild" or "free-ranging" is debatable, as there is no clear threshold defining the transition to non-wild status in non-human animals \cite{Orlando2021, Stepanoff2019}. However, no microanatomical effects of space reduction were observed in the control population compared to the captive specimens.  
%The results suggest that microanatomical variations linked to space reduction become detectable between 10 ha (control group) and 0.3 ha (enclosure group). However, current hypotheses about Neolithic husbandry practices in Western Europe, based on isotopic data indicating a more herbivorous diet in pigs, favor extensive, semi-extensive, or mixed farming systems (i.e., integrated crop-livestock systems) \cite{Navarrete2017, Balasse2014, Bogaard2007}. Thus, the space reduction observed here (10 ha vs. 0.3 ha) might not be directly applicable to early Neolithic wild boars.  
\section{Conclusion}
This preliminary study focuses on the impact of captivity on the microanatomy of this short bone, devoid of muscle insertion, in a wild ungulate (\textit{Sus scrofa}). The microanatomy of the talus in wild boars is characterized by compact bone surrounding a relatively dense trabecular bone. While the distribution of compact bone does not appear to exhibit phenotypic variations related to life in captivity, the organization of trabecular bone differs between specimens that lived in captivity and those in the wild, as the former display denser bones with a less loose trabecular network. Furthermore, variations in microanatomy are not significantly associated with either sexual dimorphism or weight. They thus appear to be clearly linked to life in captivity. Additionally, some specimens from the 
%à verifier
Late/Final Mesolithic seem to show signs of life in captivity. 
\section{Acknowledgements}

>>>>

The SVM average cross-validation score was 93.3\% and the accuracy on the final test was 84.6\%. recent/final Mesolithic specimens ARC-12990, J234, EG235, G236, C235 and C238 were classified as having microanatomy associated with captivity. Specimens E143-7, E236, G238, K135, M212, ARC-16242, RAN-270, RAN-397, RAN-398, RAN-460, RAN-545, RAN-706 were classified as having microanatomy associated with life in the wild.

@@

%verif si je la mets \subsection{Limitations}  
%As previously mentioned, the actualistic referential framework used here has several limitations. The potential reduction in rooting behavior due to supplemental feeding in the Chambord and Compiègne wild boars may have altered their locomotion. Notably, the control population already occupies a relatively restricted space compared to the Chambord and Compiègne specimens. Moreover, classifying this fenced-forest population as "wild" or "free-ranging" is debatable, as there is no clear threshold defining the transition to non-wild status in non-human animals \cite{Orlando2021, Stepanoff2019}. However, no microanatomical effects of space reduction were observed in the control population compared to the captive specimens.  
%The results suggest that microanatomical variations linked to space reduction become detectable between 10 ha (control group) and 0.3 ha (enclosure group). However, current hypotheses about Neolithic husbandry practices in Western Europe, based on isotopic data indicating a more herbivorous diet in pigs, favor extensive, semi-extensive, or mixed farming systems (i.e., integrated crop-livestock systems) \cite{Navarrete2017, Balasse2014, Bogaard2007}. Thus, the space reduction observed here (10 ha vs. 0.3 ha) might not be directly applicable to early Neolithic wild boars.  
\section{Conclusion}
This preliminary study focuses on the impact of captivity on the microanatomy of this short bone, devoid of muscle insertion, in a wild ungulate (\textit{Sus scrofa}). The microanatomy of the talus in wild boars is characterized by compact bone surrounding a relatively dense trabecular bone. While the distribution of compact bone does not appear to exhibit phenotypic variations related to life in captivity, the organization of trabecular bone differs between specimens that lived in captivity and those in the wild, as the former display denser bones with a less loose trabecular network. Furthermore, variations in microanatomy are not significantly associated with either sexual dimorphism or weight. They thus appear to be clearly linked to life in captivity. Additionally, most of the specimens from the 
%à verifier
Late/Final Mesolithic seem to show signs of life in captivity. 
\section{Acknowledgements}
